**Supplementary Information**

### 2020-02-26 submission/manuscript.tex
Results done.

All these ecological roles are carried out by a taxonomically diverse
community of microorganisms. At higher taxonomic ranks a core set of
macrophyte epiphytic taxa was described consisting mainly of
\emph{Alphaproteobacteria}, \emph{Gammaproteobacteria},
\emph{Desulfobacterota}, \emph{Bacteroidota}, \emph{Cyanobacteria},
\emph{Actinobacteriota}, \emph{Firmicutes}, \emph{Planctomycetota},
\emph{Chloroflexi} and \emph{Verrucomicrobiota} (Crump and Koch, 2008;
Tujula \emph{et al.}, 2010; Lachnit \emph{et al.}, 2011). In contrast,
at lower taxonomic ranks host specific microbial communities were
described (Lachnit \emph{et al.}, 2011; Roth-Schulze \emph{et al.},
2016). Recently, it was shown that even different morphological niches
within the same alga had a higher influence on bacterial community
variation than biogeography or environmental factors (Morrissey \emph{et
al.}, 2019). While there is high community variation between host
species is was observed that the majority of metagenomes determined
functions were conserved both between host species and individuals
(Burke \emph{et al.}, 2011a; Roth-Schulze \emph{et al.}, 2016). This
discrepancy between taxonomic and functional composition could be
explained by the lottery hypothesis. It postulates that an initial
random colonization step is performed from a set of functionally
equivalent taxonomic groups resulting in taxonomically different
epiphytic communities sharing a core set of functional genes (Burke
\emph{et al.}, 2011a; Roth-Schulze \emph{et al.}, 2016). In addition,
some of the variation in the observed data could be attributed to
different techniques used in various studies, such as different
protocols for epiphytic cell detachment and/or DNA isolation, as no
standard protocol to study epiphytic communities was established
(Ugarelli \emph{et al.}, 2019; Korlević \emph{et al.}, submitted).

The majority of studies describing macrophyte epiphytic communities did
not encompass seasonal changes (Crump and Koch, 2008; Lachnit \emph{et
al.}, 2009; Burke \emph{et al.}, 2011b; Roth-Schulze \emph{et al.},
2016; Ugarelli \emph{et al.}, 2019). In addition, if seasonal changes
were taken into account low temporal frequency and/or methodologies that
do not allow for high taxonomic resolution were used (Tujula \emph{et
al.}, 2010; Lachnit \emph{et al.}, 2011; Miranda \emph{et al.}, 2013).
In the present study we describe the seasonal dynamics of bacterial and
archaeal communities on the surfaces of the seagrass \emph{Cymodocea
nodosa} and siphonous macroalgae \emph{Caulerpa cylindracea} determined
on a mostly monthly scale. Bacterial and archaeal epiphytes were sampled
in a meadow of \emph{Cymodocea nodosa} invaded by the invasive
\emph{Caulerpa cylindracea} and in a locality of only \emph{Caulerpa
cylindracea} located in the proximity of the meadow. In addition, for
comparison the community of the surrounding seawater was characterized.

@@

Leaves of \emph{Cymodocea nodosa} were sampled in a \emph{Cymodocea
nodosa} meadow located in the proximity of the village of Funtana
(\ang{45;10;39} N, \ang{13;35;42} E). Thalli of \emph{Caulerpa
cylindracea} were sampled in the same \emph{Cymodocea nodosa} invaded
meadow in Funtana and on a locality of only \emph{Caulerpa cylindracea}
located close to the invaded meadow. Sampling of leaves and thalli was
performed approximately monthly from December 2017 to October 2018.
Leaves and thalli were collected by diving and transported to the
laboratory in containers placed on ice and filled with site seawater.
Upon arrival to the laboratory, \emph{Cymodocea nodosa} leaves were cut
into sections of 1 -- 2 \si{\cm}, while \emph{Caulerpa cylindracea}
thalli were cut into 5 -- 8 \si{\cm} long sections. Leaves and thalli
were washed three times with sterile artificial seawater (ASW) to remove
loosely attached microbial cells. Surrounding seawater was collected in
10 \si{\l} containers by diving and transported to the laboratory where
the whole container volume was filtered through a 20 \si{\um} net. The
filtrate was further sequentially filtered through 3 \si{\um} and 0.2
\si{\um} polycarbonate membrane filters (Whatman, United Kingdom) using
a peristaltic pump. Filters were briefly dried at room temperature and
stored at \num{-80} \si{\degreeCelsius}. Seawater samples were also
collected approximately monthly from July 2017 to October 2018.

@@

DNA from surfaces of \emph{Cymodocea nodosa} and \emph{Caulerpa
cylindracea} was isolated using a previously modified and adapted
protocol that allows a selective epiphytic DNA isolation (Massana
\emph{et al.}, 1997; Korlević \emph{et al.}, submitted). Briefly, leaves
and thalli are incubated in a lysis buffer and treated with lysozyme and
proteinase K. Following the incubations, the mixture containing lyzed
epiphytic cells is separated from leaves and thalli and extracted using
a phenol-chloroform procedure. Finally, the extracted DNA is
precipitated using isopropanol. DNA from seawater picoplankton was
isolated from 0.2 \si{\um} polycarbonate filters according to (Massana
\emph{et al.}, 1997) with a slight modification. Following the
phenol-chloroform extraction steps 1/10 of chilled 3 \si{\Molar} sodium
acetate (pH 5.2) was added. DNA was precipitated by adding 1 volume of
chilled isopropanol, incubating the mixtures overnight at \num{-20}
\si{\degreeCelsius} and centrifuging at 20,000 × g and 4
\si{\degreeCelsius} for 21 \si{\minute}. The pellet was washed twice
with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged
after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 5
\si{\minute}. Dried pellets were resuspended in 50 -- 100 \si{\ul} of
deionized water.

@@

Illumina MiSeq sequencing of the V4 16S rRNA region was performed as
described previously (Korlević \emph{et al.}, submitted). The V4 region
of the 16S rRNA gene was amplified using a two-step PCR procedure. In
the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R
(5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project
(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
were used to amplify the target region (Caporaso \emph{et al.}, 2012;
Apprill \emph{et al.}, 2015; Parada \emph{et al.}, 2016). These primers
contained on their 5' end a tagged sequence. Purified PCR products were
sent for Illumina MiSeq sequencing at IMGM Laboratories, Martinsried,
Germany. Before sequencing, the second PCR amplification of the two-step
PCR procedure was performed at IMGM using primers targeting the tagged
region incorporated in the first PCR. In addition, these primers
contained adapter and sample-specific index sequences. Beside samples, a
positive and negative control for each sequencing batch were sequenced.
A negative control was comprised of PCR reactions without DNA template,
while for a positive controls a mock community composed of evenly mixed
DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC,
USA) was used. The sequences obtained in this study have been submitted
to the European Nucleotide Archive (ENA) under accession numbers
\textbf{TO BE ADDED LATER!}.

@@

Obtained sequences were analyzed on the computer cluster Isabella
(University Computing Center, University of Zagreb) using mothur
(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
Standard Operating Procedure (MiSeq SOP;
\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
and recommendations given from the Riffomonas project to enhance data
reproducibility (\url{http://www.riffomonas.org/}). For alignment and
classification of sequences the SILVA SSU Ref NR 99 database (release
138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
Yilmaz \emph{et al.}, 2014). Pipeline data processing and visualization
was done using R (version 3.6.0) (R Core Team, 2019), package vegan
(version 2.5-6) (Oksanen \emph{et al.}, 2019), package tidyverse
(version 1.3.0) (Wickham \emph{et al.}, 2019) and multiple other
packages (Xie, 2014, 2015, 2019a, 2019b, 2020; Neuwirth, 2014; Xie
\emph{et al.}, 2018; Allaire \emph{et al.}, 2019; Zhu, 2019). The
detailed analysis procedure including the R Markdown file for this paper
are available as a GitHub repository (\textbf{TO BE ADDED LATER!}).
Based on the ATCC MSA-1002 mock community included in the analysis an
average sequencing error rate of 0.01 \si{\percent} was determined,
which is in line with previously reported values for next-generation
sequencing data (Kozich \emph{et al.}, 2013; Schloss \emph{et al.},
2016). In addition, the negative controls processed together with the
samples yielded on average only 2 sequences after sequence quality
curation.

@@

Chloroplast sequences were following the seasonal pattern
(\autoref{supp-chloroplast}, \autoref{pcoa}).

@@

>>>>

All these ecological roles are carried out by a taxonomically diverse
community of microorganisms. At higher taxonomic ranks a core set of
macrophyte epiphytic taxa was described consisting mainly of
\emph{Alphaproteobacteria}, \emph{Gammaproteobacteria},
\emph{Desulfobacterota}, \emph{Bacteroidota}, \emph{Cyanobacteria},
\emph{Actinobacteriota}, \emph{Firmicutes}, \emph{Planctomycetota},
\emph{Chloroflexi} and \emph{Verrucomicrobiota} (Crump and Koch, 2008;
Tujula \emph{et al.}, 2010; Lachnit \emph{et al.}, 2011). In contrast,
at lower taxonomic ranks host specific microbial communities were
described (Lachnit \emph{et al.}, 2011; Roth-Schulze \emph{et al.},
2016). Recently, it was shown that even different morphological niches
within the same alga had a higher influence on bacterial community
variation than biogeography or environmental factors (Morrissey \emph{et
al.}, 2019). While there is high community variation between host
species is was observed that the majority of metagenome determined
functions were conserved both between host species and individuals
(Burke \emph{et al.}, 2011a; Roth-Schulze \emph{et al.}, 2016). This
discrepancy between taxonomic and functional composition could be
explained by the lottery hypothesis. It postulates that an initial
random colonization step is performed from a set of functionally
equivalent taxonomic groups resulting in taxonomically different
epiphytic communities sharing a core set of functional genes (Burke
\emph{et al.}, 2011a; Roth-Schulze \emph{et al.}, 2016). In addition,
some of the variation in the observed data could be attributed to
different techniques used in various studies, such as different
protocols for epiphytic cell detachment and/or DNA isolation, as no
standard protocol to study epiphytic communities was established
(Ugarelli \emph{et al.}, 2019; Korlević \emph{et al.}, submitted).

The majority of studies describing macrophyte epiphytic communities did
not encompass seasonal changes (Crump and Koch, 2008; Lachnit \emph{et
al.}, 2009; Burke \emph{et al.}, 2011b; Roth-Schulze \emph{et al.},
2016; Ugarelli \emph{et al.}, 2019). In addition, if seasonal changes
were taken into account low temporal frequency and/or methodologies that
do not allow for high taxonomic resolution were used (Tujula \emph{et
al.}, 2010; Lachnit \emph{et al.}, 2011; Miranda \emph{et al.}, 2013).
In the present study we describe the seasonal dynamics of bacterial and
archaeal communities on the surfaces of the seagrass \emph{Cymodocea
nodosa} and siphonous macroalgae \emph{Caulerpa cylindracea} determined
on a mostly monthly scale. Bacterial and archaeal epiphytes were sampled
in a meadow of \emph{Cymodocea nodosa} invaded by the invasive
\emph{Caulerpa cylindracea} and in a locality of only \emph{Caulerpa
cylindracea} located in the proximity of the meadow. In addition, for
comparison, the community of the surrounding seawater was characterized.

@@

Leaves of \emph{Cymodocea nodosa} were sampled in a \emph{Cymodocea
nodosa} meadow located in the proximity of the village of Funtana
(\ang{45;10;39} N, \ang{13;35;42} E). Thalli of \emph{Caulerpa
cylindracea} were sampled in the same \emph{Cymodocea nodosa} invaded
meadow in Funtana and on a locality of only \emph{Caulerpa cylindracea}
located close to the invaded meadow. Sampling of leaves and thalli was
performed approximately monthly from December 2017 to October 2018
(\autoref{supp-nseq_notus}). Leaves and thalli were collected by diving
and transported to the laboratory in containers placed on ice and filled
with site seawater. Upon arrival to the laboratory, \emph{Cymodocea
nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while
\emph{Caulerpa cylindracea} thalli were cut into 5 -- 8 \si{\cm} long
sections. Leaves and thalli were washed three times with sterile
artificial seawater (ASW) to remove loosely attached microbial cells.
Surrounding seawater was collected in 10 \si{\l} containers by diving
and transported to the laboratory where the whole container volume was
filtered through a 20 \si{\um} net. The filtrate was further
sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate
membrane filters (Whatman, United Kingdom) using a peristaltic pump.
Filters were briefly dried at room temperature and stored at \num{-80}
\si{\degreeCelsius}. Seawater samples were also collected approximately
monthly from July 2017 to October 2018.

@@

DNA from surfaces of \emph{Cymodocea nodosa} and \emph{Caulerpa
cylindracea} was isolated using a previously modified and adapted
protocol that allows for a selective epiphytic DNA isolation (Massana
\emph{et al.}, 1997; Korlević \emph{et al.}, submitted). Briefly, leaves
and thalli are incubated in a lysis buffer and treated with lysozyme and
proteinase K. Following the incubations, the mixture containing lyzed
epiphytic cells is separated from leaves and thalli and extracted using
a phenol-chloroform procedure. Finally, the extracted DNA is
precipitated using isopropanol. DNA from seawater picoplankton was
isolated from 0.2 \si{\um} polycarbonate filters according to (Massana
\emph{et al.}, 1997) with a slight modification. Following the
phenol-chloroform extraction steps 1/10 of chilled 3 \si{\Molar} sodium
acetate (pH 5.2) was added. DNA was precipitated by adding 1 volume of
chilled isopropanol, incubating the mixtures overnight at \num{-20}
\si{\degreeCelsius} and centrifuging at 20,000 × g and 4
\si{\degreeCelsius} for 21 \si{\minute}. The pellet was washed twice
with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged
after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 5
\si{\minute}. Dried pellets were resuspended in 50 -- 100 \si{\ul} of
deionized water.

@@

Illumina MiSeq sequencing of the V4 16S rRNA region was performed as
described previously (Korlević \emph{et al.}, submitted). The V4 region
of the 16S rRNA gene was amplified using a two-step PCR procedure. In
the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R
(5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project
(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
were used (Caporaso \emph{et al.}, 2012; Apprill \emph{et al.}, 2015;
Parada \emph{et al.}, 2016). These primers contained on their 5' end a
tagged sequence. Purified PCR products were sent for Illumina MiSeq
sequencing at IMGM Laboratories, Martinsried, Germany. Before sequencing
at IMGM, the second PCR amplification of the two-step PCR procedure was
performed using primers targeting the tagged region incorporated in the
first PCR. In addition, these primers contained adapter and
sample-specific index sequences. Beside samples, a positive and negative
control for each sequencing batch was sequenced. Negative control was
comprised of PCR reactions without DNA template, while for a positive
control a mock community composed of evenly mixed DNA material
originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
used. The sequences obtained in this study have been submitted to the
European Nucleotide Archive (ENA) under accession numbers \textbf{TO BE
ADDED LATER!}.

@@

Obtained sequences were analyzed on the computer cluster Isabella
(University Computing Center, University of Zagreb) using mothur
(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
Standard Operating Procedure (MiSeq SOP;
\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
and recommendations given from the Riffomonas project to enhance data
reproducibility (\url{http://www.riffomonas.org/}). For alignment and
classification of sequences the SILVA SSU Ref NR 99 database (release
138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
Yilmaz \emph{et al.}, 2014). Pipeline data processing and visualization
was done using R (version 3.6.0) (R Core Team, 2019), packages vegan
(version 2.5-6) (Oksanen \emph{et al.}, 2019), and tidyverse (version
1.3.0) (Wickham \emph{et al.}, 2019) and multiple other packages (Xie,
2014, 2015, 2019a, 2019b, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018;
Allaire \emph{et al.}, 2019; Zhu, 2019). The detailed analysis procedure
including the R Markdown file for this paper are available as a GitHub
repository (\textbf{TO BE ADDED LATER!}). Based on the ATCC MSA-1002
mock community included in the analysis an average sequencing error rate
of 0.01 \si{\percent} was determined, which is in line with previously
reported values for next-generation sequencing data (Kozich \emph{et
al.}, 2013; Schloss \emph{et al.}, 2016). In addition, the negative
controls processed together with the samples yielded on average only 2
sequences after sequence quality curation.

@@

Sequencing of the 16S rRNA V4 region yielded a total of 1.8 million
sequences after quality curation and exclusion of eukaryotic,
chloroplast, mitochondrial and no relative sequences
(\autoref{supp-nseq_notus}). A total of 35 samples originating from
epiphytic archaeal and bacterial communities associated with surfaces of
the seagrass \emph{Cymodocea nodosa} and macroalga \emph{Caulerpa
cylindracea} were analyzed. In addition, 18 samples (one of the samples
was sequenced two times) originating from picoplankton archaeal and
bacterial communities in the surrounding seawater were also processed.
The number of reads per sample ranged between 8,407 and 77,465 sequences
(\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
applied the rarefaction curves did not level off that is a common
observation in high-throughput 16S rRNA amplicon sequencing approaches
(\autoref{supp-rarefaction}). Following quality curation and exclusion
of sequences mentioned before reads were clustered into 28,726 different
OTUs at a similarity level of 97 \si{\percent}. Reads numbers were
normalized to the minimum number of sequences, 8,407
(\autoref{supp-nseq_notus}), through rarefaction resulting in 16,963
different OTUs that ranged from 365 to 2,019 OTUs per sample
(\autoref{supp-calculators}). To determine seasonal changes of richness
and diversity the Observed Number of OTUs, Chao1, ACE, Exponential
Shannon (Jost, 2006) and Inverse Simpson were calculated after
normalization through rarefaction. Generally, richness estimators and
diversity indices showed similar trends. On average, higher values were
found for \emph{Caulerpa cylindracea} (invaded {[}Number of OTUs,
1,686.9 ± 135.8 OTUs{]} and noninvaded {[}Number of OTUs, 1,742.8 ±
152.4 OTUs{]}), middle vales for \emph{Cymodocea nodosa} (Number of
OTUs, 1,047.7 ± 212.1 OTUs) and lower values for picoplankton
communities in the surrounding seawater (Number of OTUs, 528.4 ± 144.5
OTUs) (\autoref{supp-calculators}). Seasonal changes did not show such
large dissimilarities. Seawater communities richness was stable during
the studied period with the exception of one sampling point in December
2017 when larger values were observed. \emph{Cymodocea nodosa}
communities showed a slow increase towards the end of the study, while
\emph{Caulerpa cylindracea} (invaded and noninvaded) communities were
characterized by slightly larger values in Spring and Summer in
comparison to Autumn and Winter (\autoref{supp-calculators}).

To determine the proportion of shared archaeal and bacterial OTUs and
communities sampled in different environments the Jaccard's Similarity
Coefficient on presence-absence data and Bray-Curtis Similarity
Coefficient were, respectively, calculated. Coefficients were determined
after normalization through rarefaction and binning of samples from a
particular environment. The highest proportion of shared OTUs and
community was found between invaded and noninvaded \emph{Caulerpa
cylindracea} (Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared
values were calculated between seawater and epiphytic communities
(\autoref{matrix}). Shared proportion between \emph{Cymodocea nodosa}
and \emph{Caulerpa cylindracea} were approximately in the middle between
these two extremes. To assess seasonal changes in the proportion of
shared OTUs and communities the Jaccard's and Bray-Curtis Similarity
Coefficients were calculated between consecutive sampling points
(\autoref{shared}). Both coefficients showed similar trends. Temporal
proportional changes were more pronounced for seawater in comparison to
\emph{Cymodocea nodosa} and especially \emph{Caulerpa cylindracea}
associated communities (\autoref{shared}). To further disentangle the
environmental and seasonal community dissimilarity a Principal
Coordinates Analysis (PCoA) based on Bray-Curtis distances and OTU
abundances was applied. It showed a clear separation between planktonic
and surface associated communities (\autoref{pcoa}). In addition, a
separation of epiphytic bacterial and archaeal communities based on host
species was determined. This separation was further supported by ANOSIM
(R = 0.96, \emph{p} \textless{} 0.001). Seasonal changes of seawater
communities indicated a separation between Spring, Summer and
Autumn/Winter samples (ANOSIM, R = 0.64, \emph{p} \textless{} 0.001).
Epiphytic microbial communities associated with \emph{Cymodocea nodosa}
showed a similar pattern (ANOSIM, R = 0.54, \emph{p} \textless{} 0.01),
while communities from the surfaces of \emph{Caulerpa cylindracea}
indicated a non so strongly supported, as in previous cases, separation
between Summer and Autumn/Winter/Spring samples (ANOSIM, R = 0.30,
\emph{p} \textless{} 0.01) (\autoref{pcoa}).

The microbial taxonomic composition of both, macrophyte associated and
seawater communities, was dominated by \emph{Bacteria} (99.1 ± 2.1
\si{\percent}) over \emph{Archaea} (0.9 ± 2.1 \si{\percent})
(\autoref{community}). Three phyla were characteristic for
\emph{Archaea}: \emph{Crenarchaeota}, \emph{Nanoarchaeota} and
\emph{Thermoplasmatota}. Their contribution was larger only in seawater
communities during December 2017 (\autoref{community}). Higher relative
abundances of chloroplast related sequences were only observed in
surface associated communities, with higher values in Autumn/Winter
(37.2 ± 11.2 \si{\percent}) in comparison to Spring/Summer (20.9 ± 9.7
\si{\percent}) (\autoref{supp-chloroplast}). Generally, at higher
taxonomic ranks (phylum-class) epiphytic and seawater microbial
communities were composed of similar bacterial taxa. Seawater
communities were mainly comprised of \emph{Actinobacteriota},
\emph{Bacteroidota}, \emph{Cyanobacteria}, \emph{Alphaproteobacteria},
\emph{Gammaproteobacteria} and \emph{Verrucomicrobiota}. Communities
associated with \emph{Cymodocea nodosa} were consisted of same groups
with the addition of \emph{Planctomycetota} whose contribution was
higher in summer 2018. In addition, communities from invaded and
noninvaded \emph{Caulerpa cylindracea} were similar and characterized by
same groups as seawater and \emph{Cymodocea nodosa} communities with the
addition of \emph{Desulfobacterota} (\autoref{community}). Larger
differences between environments and host species could be observed at
lower taxonomic ranks (\autoref{cyano} -- \ref{desulfo}).

\emph{Cyanobacteria} related sequences were comprising, on average, 5.5
± 4.4 \si{\percent} of total sequences (\autoref{cyano}). Higher
proportions were found for \emph{Cymodocea nodosa} (16.4 ± 5.3
\si{\percent}) and \emph{Caulerpa cylindracea} (invaded {[} (7.7 ± 3.9
\si{\percent}){]} and noninvaded {[} (7.8 ± 2.4 \si{\percent}){]})
associated communities in autumn and for seawater communities in winter
(8.8 ± 7.4 \si{\percent}). Large taxonomic differences between surface
associated and seawater cyanobacterial communities were observed.
Seawater communities were mainly comprised of \emph{Cyanobium} and
\emph{Synechococcus}, while surface associated communities were
consisted of and sequences without known relatives within
\emph{Cyanobacteriales} and \emph{Cyanobacteriia} (\autoref{cyano}). In
addition, seasonal changes in surface associated communities were
observed with \emph{Cyanobacteriales} and \emph{Cyanobacteriia} no
relative sequences comprising larger proportions in autumn and winter
and \emph{Acrophormium}, \emph{Phormidesmis} and no relative
\emph{Nodosilineaceae} and in spring and summer (\autoref{cyano}).

Sequences classified as \emph{Bacteroidota} were comprising, on average,
19.2 ± 5.5 \si{\percent} of all sequences (\autoref{bactero}). Similarly
to \emph{Cyanobacteria}, large differences in the taxonomic composition
between seawater and surface associated communities were found
(\autoref{bactero}). The seawater community was characterized by the NS4
and NS5 marine groups, uncultured \emph{Cryomorphaceae}, uncultured
\emph{Flavobacteriaceae}, NS11-12 marine group, \emph{Balneola},
uncultured \emph{Balneolaceae} and \emph{Formosa}. In contrast, surface
associated communities were characterized by \emph{Lewinella},
\emph{Portibacter}, \emph{Rubidimonas}, no relative
\emph{Saprospiraceae}, uncultured \emph{Saprospiraceae}, no relative
\emph{Flavobacteriaceae} and uncultured \emph{Rhodothermaceae}. Surface
associated \emph{Bacteroidota} communities were very diverse as could be
observed in the the high proportion of taxa that grouped as other
\emph{Bacteroidota} (\autoref{bactero}).

On average, \emph{Alphaproteobacteria} were in comparison to other high
rank taxa the largest taxonomic group, comprising 29.2 ± 12.0
\si{\percent} of all sequences (\autoref{alpha}). In accordance to
previous taxa, high differences between seawater and surface associated
communities were observed. Picoplankton communities were composed mainly
of the SAR11 clade, AEGEAN-169 marine group, SAR116 clade, no relative
\emph{Rhodobacteraceae}, HIMB11 and OCS116 clade, while surface
associated communities were composed in high proportion of no relative
\emph{Rhodobacteraceae} and to a lesser degree of \emph{Pseudoahrensia},
no relative \emph{Alphaproteobacteria} and no relative
\emph{Hyphomonadaceae}. Representatives of no relative
\emph{Rhodobacteraceae} were comprising in on average 40.6 ± 23.2
\si{\percent} of all alphaproteobacterial sequences from the epiphytic
community (\autoref{alpha}).

Sequences related to \emph{Gammaproteobacteria} were comprising, on
average, 18.7 ± 3.9 \si{\percent} of all sequences (\autoref{gamma}).
Similarly to previous taxa, large taxonomic differences between seawater
and surface associated communities were found. Seawater communities were
mainly comprised of the OM60 (NOR5) clade, \emph{Litoricola},
\emph{Acinetobacter} and the SAR86 clade, while epiphytic communities
were mainly composed of no relative \emph{Gammaproteobacteria} and
\emph{Granulosicocus}. Beside these two groups specific to all three
epiphytic communities, \emph{Cymodocea nodosa} was characterized by
\emph{Arenicella}, no relative \emph{Burkholderiales} and
\emph{Methylotenera}, while \emph{Thioploca}, no relative
\emph{Cellvibrionaceae} and \emph{Reinekea} were more specific to both
invaded and noninvaded \emph{Caulerpa cylindracea}. In addition,
similarly to \emph{Bacteroidota}, a large proportion of the surface
associated community was grouped as other \emph{Gammaproteobacteria}
indicating high diversity within this group (\autoref{gamma}).

In contrast to previously described high rank taxa,
\emph{Desulfobacterota} were specific to \emph{Caulerpa cylindracea}. On
average they were comprising 11.2 ± 13.3 \si{\percent} of all sequences.
While seawater and \emph{Cymodocea nodosa} communities were consisted of
only 0.1 ± 0.08 \si{\percent} and 1.0 ± 0.7 \si{\percent}
\emph{Desulfobacterota} sequences, respectively, in the invaded and
noninvaded \emph{Caulerpa cylindracea} communities their proportion was
25.7 ± 11.3 \si{\percent} and 24.0 ± 4.3 \si{\percent}, respectively
(\autoref{desulfo}). In addition, \emph{Caulerpa cylindracea} associated
communities were characterized by higer proportions in Winter and Summer
(invaded, 30.9 ± 12.4 \si{\percent}; noninvaded, 26.9 ± 3.4
\si{\percent}) in comparison to Autumn and Spring (invaded, 22.8 ± 10.3
\si{\percent}; noninvaded, 21.9 ± 3.8 \si{\percent}). The community was
mainly consisted of no relative \emph{Desulfobacteraceae},
\emph{Desulfatitalea}, no relative \emph{Desulfobulbaceae},
\emph{Desulfobulbus}, no relative \emph{Desulfocapsaceae},
\emph{Desulfopila}, \emph{Desulforhopalus}, \emph{Desulfotalea},
SEEP-SRB4 and uncultured \emph{Desulfocapsaceae} (\autoref{desulfo}).

@@

{\centering \includegraphics[width=0.7\linewidth]{../results/figures/matrix}

\caption{Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between communties associated with the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and coomunities in the surrounding seawater.\label{matrix}}\label{fig:unnamed-chunk-1}
\end{figure}

### 2020-02-26 submission/supplementary.Rmd
Results done.

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE}
opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)
opts_chunk$set("message" = FALSE)

@@

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=1, nsmall=1, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
knitr::knit_hooks$set(inline=inline_hook)
```

@@

>>>>

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE, message=FALSE}
opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)
opts_chunk$set("message" = FALSE)

@@

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=1, nsmall=1, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
knitr::knit_hooks$set(inline=inline_hook)

library("kableExtra")
```

@@

## Supplementary Table

```{r nseq_notus}
# Loading OTU/sample data
shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")

# Calculating the number of sequences and OTUs in each sample
nseq_notus <- shared %>%
  mutate(nseq=rowSums(select(., starts_with("Otu")))) %>%
  mutate(notus=rowSums(select(., starts_with("Otu"))!=0)) %>%
  select(Group, nseq, notus)

# Loading metadata 
metadata <- read_tsv("data/raw/metadata.csv")

# Joining metadata with the number of sequences and OTUs
invisible(Sys.setlocale(locale="en_GB.utf8"))
metadata_nseq_notus <- inner_join(metadata, nseq_notus, by=c("ID"="Group")) %>%
  select(ID, station, date, season, nseq, notus) %>%
  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
  mutate(date=format(date, "%d %B %Y")) %>%
  mutate(date=str_replace(date, "^0", "")) %>%
  arrange(match(station, c("F", "FCyM", "FCaM", "FCa"))) %>%
  mutate(station=str_replace(station, "^F$", "Seawater")) %>%
  mutate(station=str_replace(station, "^FCyM$", "\\\\textit{Cymodocea nodosa} (Invaded)")) %>%
  mutate(station=str_replace(station, "^FCaM$", "\\\\textit{Caulerpa cylindracea} (Invaded)")) %>%
  mutate(station=str_replace(station, "^FCa$", "\\\\textit{Caulerpa cylindracea} (Nonnvaded)")) %>%
  mutate(ID=str_replace(ID, "^23_1$", "23a")) %>%
  mutate(ID=str_replace(ID, "^23_2$", "23b")) %>%
  mutate(nseq=format(nseq, big.mark=","), notus=format(notus, big.mark=","))

# Generating the table
kable(metadata_nseq_notus, format="latex", booktabs=TRUE, linesep="", longtable=TRUE, escape=FALSE, col.names=c("Sample ID", "Community Type", "Date", "Season", "No. of Sequences", "No. of OTUs"), caption="Sample ID, Community Type, Sampling Date and Season, No. of Sequences and No. of OTUs of each sample.\\label{nseq_notus}", align="c") %>%
  kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9) %>%
  row_spec(0, bold=TRUE) %>%
  column_spec(1, width="6em")
```
